### 2013-07-08 paper.tex
Doing commits manually on Github now, git cli is just screwing around.

Unfortunately estimating tree volume is expensive, time consuming and in most cases the data collected doesn't meet today's standards \cite{digital imaged based tree measurement for forest inventory}
Most methods used today to measure individual trees are often done manually with a variety of tools. However the use of these analogue approaches does not come without it's downfalls.

@@

The next stage is retrieving the real world height to each set of pixel coordinates as well as the width between them, which will be the diameter of the tree. The first step taken is to get the height. The vertical image plate scale, which is the ratio between degrees and pixels in the vertical axis, is calculated by dividing the vertical resolution of the image with the vertical angle of view. This can be done because it is assumed that the silhouette of the tree is on a plane in the real world which corresponds to the image plane on the camera sensor. With this method it is possible to find out the angle to each pixel coordinate, for example $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view. This is angle is denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}. However if the pixel coordinate was identified in the bole image, when the phone was tilted upwards, then angle between the distance to the tree and the lower boundary of the vertical angle of view, $B$, will need to be added, which is represented by $\varphi$ in figure \ref{triangle2}. The diagonal distance to $P$ is also calculated so that the width can later be determined.

@@

\section{Discussion}
\subsection{Error tolerance}
As seen in table \ref{heights} either when the vertical tilt angle is very wide, standing close to the tree, or very narrow, standing far away, leads to error. It's not clear what is causing this but there a two possible reasons. The first one is that because of the short focal length of the camera and overall poor quality when compared to a real camera, causes various forms of distrotion, especially around the edges of the image. When the person is standing close to the tree so the vertical tilt angle is over 45\degree, the points tend to lie exactly there, around the edges of the image. The distortion then affects these points causing the pixel measurements to be incorrect. The other problem occurs in the direct opposite situation, when the person is standing far away and the angle is very narrow. Naturally what happens is that the resolution of the camera just isn't high enough and the pixel measurements just aren't accurate anymore. However most of this distortion seems to only affect the vertical measurements because when looking at table \ref{widths} it is clearly seen that the readings do not stray much at all from their actual values. Despite this the deviation in the height readings still alters the volume calculations. Ideally a person should stand so that the vertical tilt angle lies between 45\degree and 25\degree which basically means they should stand somewhere between one tree length to two tree length's from the actual tree.
\subsection{Limitations}
As pointed out in the previous section the main hindrance attatched to this application is the actual camera. The problem is that most mobile phone cameras have a lot of different distortions that make it hard to make accurate measurements. Also when the tree is far away, the resolution of the camera can have a large affect on the final result. Human error is still large factor. If the person is not holding the phone properly, takes a blurry photo or in any other way alters the image, will cause the calculations to be incorrect. Mechanical Turk is also a weak link because it is essentially humans doing the work and there is nothing stopping them from also making mistakes. 
\subsection{Future development}
The main area that could be expanded on is the edge detection. As explained in the background, today computer vision is still far from perfect. However that will hopefully change in the near future and things like Mechanical Turk can be dropped for a more efficient system that produces little to no error and works autonomously. If camera technology in mobile phones doesn't improve then being able to correct the various distortions would also make an improvement. The possibilty might also exist in the future to have the application calculate extra things like canopy area and even adding in the volume of external branches to the main volume.

>>>>

Unfortunately estimating tree volume is expensive, time consuming and in most cases the data collected doesn't meet today's standards \cite{digital imaged based tree measurement for forest inventory}.
Most methods used today to measure individual trees are often done manually with a variety of tools. However the use of these analogue approaches does not come without it's downfalls.

@@

The next stage is retrieving the real world height to each set of pixel coordinates as well as the width between them, which will be the diameter of the tree. The first step taken is to get the height. The vertical image plate scale, which is the ratio between degrees and pixels in the vertical axis, is calculated by dividing the vertical resolution of the image with the vertical angle of view. This can be done because it is assumed that the silhouette of the tree is on a plane in the real world which corresponds to the image plane on the camera sensor. With this method it is possible to find out the angle to each pixel coordinate, for example $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view. This is angle is denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}. However if the pixel coordinate was identified in the bole image, when the phone was tilted upwards, then angle between the distance to the tree and the lower boundary of the vertical angle of view, $B$, will need to be added, which is represented by $\varphi$ in figure \ref{triangle2}. To calculate the real world height to each set of pixel coordinates, the horizontal distance to the tree, $d$ in figure \ref{horizontal_triangle}, is multiplied wtih the angle to the pixel coordinate. In the DBH picture this angle is just $\theta$ and in the bole picture it is $\theta + \varphi$. The diagonal distance to $P$ is also calculated so that the width can later be determined.

@@

\section{Discussion}
\subsection{Error tolerance}
As seen in table \ref{heights} either when the vertical tilt angle is very wide, standing close to the tree, or very narrow, standing far away, leads to error. It's not clear what is causing this but there a two possible reasons. The first one is that because of the short focal length of the camera and overall poor quality when compared to a real camera, causes various forms of distrotion, especially around the edges of the image. When the person is standing close to the tree so the vertical tilt angle is over 45\degree, the points tend to lie exactly there, around the edges of the image. The distortion then affects these points causing the pixel measurements to be incorrect. The other problem occurs in the direct opposite situation, when the person is standing far away and the angle is very narrow. Naturally what happens is that the resolution of the camera just isn't high enough and the pixel measurements just aren't accurate anymore. However most of this distortion seems to only affect the vertical measurements because when looking at table \ref{widths} it is clearly seen that the readings do not stray much at all from their actual values. Despite this the deviation in the height readings still alters the volume calculations, as observed in table \ref{volumes}. Ideally a person should stand so that the vertical tilt angle lies between 45\degree and 25\degree which basically means they should stand somewhere between one tree length to two tree length's from the actual tree.
\subsection{Limitations}
As pointed out in the previous section the main hindrance attatched to this application is the actual camera. The problem is that most mobile phone cameras have a lot of different distortions that make it hard to make accurate measurements. Also when the tree is far away, the resolution of the camera can have a large affect on the final result. Human error is still large factor. If the person is not holding the phone properly, takes a blurry photo or in any other way alters the image, will cause the calculations to be incorrect. Mechanical Turk is also a weak link because it is essentially humans doing the work and there is nothing stopping them from also making mistakes. Although the application has removed a lot of possible human error it has also introduced new machine error. Sensor data from the phone can be inaccurate. The built-in gyroscope and accelerometer can sometimes give very rough measurements and a few degrees can change the whole outcome. Plus if they are not callibrate properly then that will also lead to error.
\subsection{Future development}
The main area that could be expanded on is the edge detection. As explained in the background, today computer vision is still far from perfect. However that will hopefully change in the near future and things like Mechanical Turk can be dropped for a more efficient system that produces little to no error and works autonomously. If camera technology in mobile phones doesn't improve then being able to correct the various distortions would also make an improvement. The possibilty might also exist in the future to have the application calculate extra things like canopy area and volume of external branches and limbs.
